Analysis of competitive surfing tournaments with generalized Bradley-Terry likelihoods

### 2022-04-23 inst/surfing_jsa_rebuttal_referee2.tex

\textcolor{blue}{These comments are entirely reasonable and, on
  re-reading the manuscript, we agree that we could have done a much
  better job describing the mathematical and statistical basis for our
  conclusions, especially the motivation for the likelihood equations.
  In the resubmission we have added a large amount of explanation that
  hopefully makes our reasoning more clear.\\ \\
  However, we observe that {\em all} our calculations are available
  for public inspection on gihub at our repo {\tt RobinHankin/hyper2}.
  In particular, file {\tt inst/surfing.Rmd} can reproduce all our
  findings.  However, the nitty-gritty detail of our R idiom is, we
  suggest, not really suitable for a JSA manuscript.  We believe that
  our enhanced reporting in the revision is a good compromise between
  not enough detail and too much.
  \begin{itemize}
  \item We have provided more detail (mathematical and computational)
    for the Fisher information argument for retaining the data with
    three surfers, as opposed to retaining just the pairs
  \item Added detail on the computational infrastructure for locating
    maximum likelihood estimates
  \item Referenced the computational method [{\tt equalp.test()} for
    testing the null of equal competitive abilities
  \item Introduced a new table that explicitly sets out the likelihood
    function for the reified entity Brazilian beachtype reified entity
  \end{itemize}
}

The figures and tables are poorly presented and would benefit from
significant rework in terms of formatting.

Section 7, while an interesting topic, feels like it falls more into
the realm of speculation than concrete analysis in the way it is
presented. The methodology is not very clear and it would be nice to
see a more explicit definition and derivation of pc

\textcolor{blue}{We agree that this section is more speculative than
  the remainder of the manuscript.  However, we would reiterate that
  the topic is clearly interesting and worth investigating, as both
  referees note.  Non-competitive behaviour of the sort discussed here
  is notoriously difficult to study; and in particular, formal
  statistical analysis of non-competitive behaviour is almost
  non-existant in the literature.  We are not aware of any other
  probability model that is able to even establish the existence of
  non-competitive behaviour, let alone quantifiy it.  We believe that
  our approach is new, and our probability model for non-competitive
  behaviour appropriate and plausible.\\ But, on re-reading the
  section, we agree that the presentation is indeed somewhat obscure.
  We have extensively re-written the section and believe it is much
  improved.  We include a new table motivating and discussing
  non-competitive behaviour using a particular example from the
  tournament.  We note that we presented a likelihood function for the
  three-competitor case in the first draft (it is now Table 3), and we
  consider this to be as clear as reasonably practicable.}

Overall, it feels like the main conclusion is ``not all surfers have
the same skill". This is not insignificant, however, it is not
particularly surprising and is one that could have been reached using
a number of different methods. Had the authors done a more complete
job of fleshing out why the Bradley-Terry approach was better for
reaching this conclusion than others, I would be more satisfied, but
as it stands it falls a little flat.

### 2022-04-23 inst/surfing_jsa.tex
"speculation [rather] than concrete analysis" referee's comment

\documentclass{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[pdftex]{hyperref}

@@

\begin{table}[h]
\begin{tabular}{l|l|l}
wave type  & $\operatorname{Prob}(p_1\succ p_2)$ & $\operatorname{Prob}(p_2\succ p_1)$\\ \hline
beach & $\frac{p_1+B}{p_1+p_2+B}$ & $\frac{p_2}{p_1+p_2+B}$\\
point & $\frac{p_1+B}{p_1+p_2+B}$ & $\frac{p_2}{p_1+p_2+B}$\\
reef  & $\frac{p_1\vphantom{B}}{p_1+p_2}$ & $\frac{p_2}{p_1+p_2}$\\
\end{tabular}
\caption{Likelihood function for results between surfer 1 (Brazilian)
  and surfer 2 (non-Brazilian) with \label{probabilitybrazilian} Bradley-Terry
  strengths $p_1,p_2$ respectively, and a ``Brazilian wavetype" entity with
  strength $B$ that quantifies the enhanced Brazilian performance on reef wave types}
\end{table}

>>>>

\documentclass{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{diagbox}
\usepackage[pdftex]{hyperref}

@@

\begin{table}[h]
\begin{tabular}{l|l|l}
wave type  & $\operatorname{Prob}(p_1\succ p_2)$ & $\operatorname{Prob}(p_2\succ p_1)$\\ \hline
beach & $\frac{p_1+B}{p_1+p_2+B}$ & $\frac{p_2}{p_1+p_2+B}$\\
point & $\frac{p_1+B}{p_1+p_2+B}$ & $\frac{p_2}{p_1+p_2+B}$\\
reef  & $\frac{p_1\vphantom{B}}{p_1+p_2}$ & $\frac{p_2}{p_1+p_2}$\\
\end{tabular}
\caption{Likelihood function for results between surfer 1 (Brazilian)
  and surfer 2 (non-Brazilian) with \label{probabilitybrazilian} Bradley-Terry
  strengths $p_1,p_2$ respectively, and a ``Brazilian wavetype" entity with
  strength $B$ that quantifies the enhanced Brazilian performance on reef wave types}
\end{table}
\clearpage
\newpage

\begin{table}[h]
  \begin{tabular}{l|l|l|l}\\
    \backslashbox{order}{value}          & B & F & S\\ \hline
    $\mbox{B}\succ\mbox{F}\succ\mbox{S}$ & 1 & 0 & 0\\
    $\mbox{B}\succ\mbox{S}\succ\mbox{F}$ & 1 & 0 & 1\\
    $\mbox{F}\succ\mbox{B}\succ\mbox{S}$ & 0 & 1 & 0\\
    $\mbox{F}\succ\mbox{S}\succ\mbox{B}$ & 0 & 1 & 1\\
    $\mbox{S}\succ\mbox{B}\succ\mbox{F}$ & 1 & 0 & 2\\
    $\mbox{S}\succ\mbox{F}\succ\mbox{B}$ & 0 & 1 & 2\\
  \end{tabular}
  \caption{Simplified payoff table\label{valuetable} for Buchan (B),
    Freestone (F), and Smith (S).  Buchan and Freestone, being
    Australian, have a payoff that reflects their likelihood of being
    picked for the national Olympic squad: they are interested in
    whether or not they beat their countryman.  Thus Buchan's payoff
    shows whether or not he beat Freestone, and vice versa.  Smith,
    being South African, is motivated by his ranking against his
    competitors and here we show Smith with a Borda-type payoff.
    Smith observes that {\em both} Buchan and Freestone are
    indifferent between outcomes $\left\lbrace
    \mbox{B}\succ\mbox{F}\succ\mbox{S},
    \mbox{B}\succ\mbox{S}\succ\mbox{F},
    \mbox{S}\succ\mbox{B}\succ\mbox{F}\right\rbrace$, and may be able
    to exploit this indifference by strategies increasing the
    probability of his desired outcome,
    $\mbox{S}\succ\mbox{B}\succ\mbox{F}$}
\end{table}
